---
title: "Bayesian method for analyzing results of acceptability judgment experiment"
output: html_document
---

### Required packages
```{r}
library(tidyverse)
library(lmerTest)
library(brms)
library(bayestestR)

options(dplyr.summarise.inform = FALSE)
```

### Read data
```{r}
data = read.csv("./fakedata_bayesian.csv")
```

### Visualize data
The data has typical 2x2 design, manipulating the type of structure in a sentence (non-island|island) and whether there's an extraction out of the structure (+|-).
```{r}
data_summary = data %>% group_by(structure, extraction) %>%
  summarize(n = n(),
            mu = mean(judgments),
            sigma = sd(judgments),
            SE = sigma/sqrt(n))

data_summary$extraction = as.factor(data_summary$extraction)
```

```{r}
data_summary %>%
  ggplot(aes(x=extraction, y=mu))+
  geom_point()+
  geom_path(aes(group = structure, linetype = structure))+
  geom_errorbar(aes(ymin=mu-SE, ymax=mu+SE), width=0.1)+
  expand_limits(y = c(-1, 1))+
  theme_bw()
```

### Frequentist data analysis using `lmer`
Note that the model isn't maximally random as such a model didn't converge.
```{r}
model_lmer = lmer(judgments ~ structure*extraction + (1+extraction | subject) + (1 | item), data=data)
summary(model_lmer)
```

### Bayesian data analysis
### Check default priors
We can see that the priors for class "b" (regression coefficient) are flat by default.
```{r}
default_priors = get_prior(judgments ~ structure*extraction + (1+structure*extraction | subject) + (1+structure*extraction | item), data=data)
default_priors
```

### Set priors for class "b"
Feel free to change the prior for the regression coefficients, as well as the other parameters.
```{r}
priors = c(
  set_prior("student_t(3, 0, 2.5)", class = "b")
)
```

### Fit a model
It will take a while depending on the number of chains and iterations.
```{r}
model_bayes = brm(judgments ~ structure*extraction + (1+structure*extraction | subject) + (1+structure*extraction | item), 
                  data = data,
                  prior = priors,
                  family = gaussian(),
                  chains=3, iter=3000, warmup=600,
                  save_all_pars = TRUE)

summary(model_bayes)
```

### Calculate Bayes Factor
```{r}
bayesfactor_parameters(model_bayes, null = 0)
```
